### 2016-02-15 paper/cover_letter3.tex
added prcc sensitivity and edits

We are pleased to resubmit the revised manuscript, ``A spatial model of the
efficiency of T cell search in the influenza-infected lung" for JTB's 
consideration. This is a resubmission of manuscript number JTB-D-15-00501R1,
received by us on October 26, 2015 with an invitation for resubmission within 
two months.

We are happy to learn that the major concerns of Reviewer 2 and the previous 
concerns of Reviewer 1 have been addressed. Reviewer 1 now asks for alternative
models (major point 1) to achieve higher confidence in our estimation of specific
parameters. We have addressed this concern in our current response and we 
note that two separate sensitivity analyses confirmed that the parameters in 
question (T cell production rate and the chemokine secretion rate) do not have 
a strong effect on the model results. Reviewer 1 has also raised several new 
interesting questions and suggested further extensions to our model, including
new parameters describing the innate immune response (major point 4) and 
environmental factors such as vascular structure (major point 3).  We feel that these 
requests point to questions that are outside the scope of our current study, 
which focuses on comparing the effectiveness of T cells at controlling infection
in three specific influenza strains.

Our revised manuscript discusses how the model takes these additional factors
into account, but we have not explicitly added new parameters to our already 
complicated model. We anticipate several follow-on projects to explore these 
questions using new data sets and alternative models.

>>>>

We are pleased to submit the revised manuscript, ``A spatial model of the
efficiency of T cell search in the influenza-infected lung" for JTB's 
consideration. This is a resubmission of manuscript number JTB-D-15-00501R2,
received by us on December 16, 2015 with an invitation for resubmission within 
sixty days.

We are grateful for the opportunity to (re)respond to Reviewer 1's
comments. We regret that Reviewer 1 was not satisfied by our
earlier responses, but we believe that we have addressed all of
his/her concerns.  In our response, we have provided additional evidence and discussion
specifically to address Reviewer 1's comments.  We appreciate the time
spent by Reviewer 1 on the manuscript, but we respectfully disagree
with his/her opinion that we did not understand or address his/her
concerns.   At this point, we believe that we 
have adequately addressed all comments from both Reviewer 1 and 2.

%In summary, we agree with the statements and implications in Points 1, 2, 4 and 6.
%We have answered the questions in Points 3 and 5. We include detailed responses 
%to these points, but we respectfully submit that nowhere in the review is there a 
%reference to an incorrect or incomplete analysis.

### 2016-02-15 paper/response3.tex
added prcc sensitivity and edits

We are grateful for the opportunity to respond to Reviewer 1's comments. We regret that Reviewer 1 did not ultimately agree with our rebuttal, but we believe that we have directly addressed all of his/her concerns in the previous rebuttal. We are providing additional evidence and discussion as requested by the last round of revisions. While we are grateful for the time given by Reviewer 1 to review our manuscript, we respectfully disagree with his/her contention that we did not understand or address his/her concerns. We provide additional rebuttals to Reviewer 1's comments below. We believe that we have adequately addressed all comments from both Reviewer 1 and 2.

@@

Reviewer 1: The authors misinterpreted what my comments meant, and this is third time.

@@

There is no known literature with chemokine secretion rate estimations.  In this paper, we present new empirical data of \textit{in vitro} chemokine levels and extend a previously published ODE model to estimate chemokine production rates.  We are as surprised as Reviewer 1 that these production rates do not have a strong effect on our agent-based model results as determined by our sensitivity analyses.  Nevertheless, because our agent-based model examines chemokine interactions, we require a non-zero value for the production rate.  We believe that fitting the extended ODE model to our new empirical data is the best approach available to determine biologically plausible values.  Our paper contains a subsection, \textit{Models for Parameter Estimation}, that details the exact process used to estimate the chemokine production rates (page 6).  We fit the system of differential equations to empirical data using a genetic algorithm to minimize the log squared error between the model and the data (the genetic algorithm performed favorably when compared to standard Levenberg-Marquardt nonlinear regression techniques).

We have added a figure displaying the model fits to the supplement (Fig. S1, also below) and Table S3 showing the root mean squared error of the fits for completeness.

@@

\begin{response}
All biological data have a foundation using \textit{in vitro} research. We agree with Reviewer 1 that while this raises the potential caveat that some of the values that we use to parameterize our models using \textit{in vitro} generated data may not reflect \textit{in vivo} values, we, along with most other researchers in the field, use \textit{in vitro} generated values as a starting point. We explicitly state this in the manuscript. This is precisely the reason that we undertook an extensive sensitivity analysis as we acknowledge that the values we use to parameterize the model may vary from initial starting values. We do not claim that the values we estimate are absolutely reflective of \textit{in vivo} values. Instead, we use best fit estimates that are currently available as a starting point, and use sensitivity analysis to confirm the role of individual parameters on the model.​
\end{response}

@@

Chemokine does have a higher diffusion rate than virus in our model (Table 2).  It is important to note that our discussion of the infection outpacing the chemokine is different than the direct comparison between the viral diffusion rate and the chemokine diffusion rate.  Because T cells climb the chemokine gradient, a chemokine's effect is seen in the location of its local maxima, while the viral effect is seen in the spread and concentration of the virions themselves.  We have edited the text to be more clear of this distinction (\textit{Results: Spatial Effects}, page 10).

@@

To complement our spatial analysis we performed multiple sensitivity analyses on our model's free parameters.  In doing so, we examined scenarios where the chemokine's diffusion rate was increased by up to two orders of magnitude (Table 2).  Each sensitivity analysis showed that the chemokine diffusion rate was an insensitive parameter in our model (Table 3 and Supplement)

@@

We agree with the reviewer that the T cell secretion rate is an insensitive parameter in our model as determined by our multiple sensitivity analyses.  Because our agent-based model investigates the recruitment of T cells to the focus of infection, we required a non-zero secretion rate to introduce these T cells into the model.  Similar to the chemokine secretion rate as discussed in the first point and in our paper (\textit{Models for Parameter Estimation}, page 7), we fit an ODE model to empirical data to best estimate a biologically plausible value for the T cell secretion rate for use in our agent-based model.

We also agree that the recruitment of T cells to the focus of the infection is important.  We discuss this in the \textit{Spatial Effects} subsection (page 9-10) and in the discussion (page 15) as follows:

@@

The term `viral control' has been used in three different contexts.  We have explicitly addressed each point in previous responses.  In summary:

\begin{enumerate}
\item \textit{You ignored my request to investigate if virus control can be good by T cells (irrespectively of the type of virus) if chemokines diffuse faster than virus particles. Please do the simulations, and rephrase your discussion depending on the results you find.}
\begin{displayquote}
We have performed these simulations numerous times in the form of our sensitivity analyses.  Further, you seem to be misunderstanding the distinction between the differing roles of viral diffusion and chemokine diffusion.  Please refer to Response 3 of this document for more information.
\end{displayquote}

@@

\begin{displayquote}
In summary, our paper makes no attempt to model the immune response in its entirety.  Our model does represent viral
spread using parameters which likely reflect innate control mechanisms (virus secretion, infectivity, diffusion
rate, and decay rate, see Tables 1 and 2). We appreciate the reviewer’s perspective, but our goal in this
paper is to study the effect of T cells and not to fully represent the interaction between influenza strains
and the entirety of the immune response (Introduction 3rd paragraph shown below). \\

\textit{``Specifically, we focus on the interactions between activated antigen-specific CD8 T cells, cy-
tokines, and replicating influenza virus. ... Therefore, instead of developing a comprehensive
immune system model, we present a spatially explicit agent-based model (ABM) to describe T
cell interactions with chemotactic signals and a dynamically growing plaque." (page 1)} \\

We feel that careful investigation of the role of early innate control, with its many complex mechanisms, in
different influenza strains is best left to future studies.

\end{displayquote}

@@

\begin{response}
We have reviewed Oreskes et. al.  The paper correctly points out that scientists often use terms incorrectly when describing their model's veracity which leads to incorrect claims and conclusions.  Specifically, Oreskes et. al. examines four terms: `verification', `validation', `calibration', and `confirmation'.
\begin{itemize}
\item \textbf{Verification:} Oreskes et. al. states that it is not possible to verify a numerical model by comparing it to an analytical model.  We use the term `verification' one time (page 15, concluding paragraph) in the context of obtaining empirical data only (emphasis added).
\begin{displayquote}
``Empirical \textbf{verification} of the model's sensitive parameters (viral response to IgM, infectivity, viral decay rate, viral diffusion rate, and viral production rate) will be valuable to future studies."
\end{displayquote}
\item \textbf{Validation:} We do not use the term in our paper.
\item \textbf{Calibration:} Oreskes et. al. warn of the dangers of using the calibration of free parameters to emprical data as a basis for claiming model validity.  We use the term `calibration' only once (Abstract) and make no erroneous claims (emphasis added).
\begin{displayquote}
``We \textbf{calibrate} the model using viral and chemokine secretion rates we measure \textit{in vitro} together with values taken from literature."
\end{displayquote}
\item \textbf{Confirmation:} Oreskes et. al. states that observations consistent to a theory can `confirm' a theory, but warn that confirmation cannot lead to verification.  We use the term `confirm' twice and make no claims of verification or validity (emphasis added).

@@

At no point have we made claims inconsistent with Oreskes et. al.  We agree that models can inform research and use our model to propose a hypothesis of how additional T cells are limited in their effectiveness due to the spatial effects of the chemokine gradient and T cell chemotaxis.
\end{response}

>>>>

We are grateful for the opportunity to (re)respond to Reviewer 1's
comments. We regret that Reviewer 1 was not satisfied by our
earlier responses, but we believe that we directly addressed all of
his/her concerns.  Here we provide additional evidence and discussion
to respond to Reviewer 1's comments.  We appreciate the time
spent by Reviewer 1 on the manuscript, but we respectfully disagree
with his/her opinion that we did not understand or address his/her
concerns.  At this point, we believe that we have adequately addressed
all comments from both Reviewer 1 and 2.

In summary, we agree with the statements and implications in Points 1,
2, 4 and 6. We have answered the questions in Points 3 and 5. Below we
include detailed responses to these points, but we respectfully submit
that nowhere in the review is there a reference to an incorrect
or incomplete analysis.

@@

%\textbf{Reviewer 1: The authors misinterpreted what my comments meant, and this is third time.}

@@

This comment seems contradictory: On the one hand it asks for more
details about how we estimate chemokine production rate, and on the
other hand suggests that chemokine production rate doesn't need to
be estimated because it does not impact clearance.  Our analysis
agrees with his statement that it does not impact clearance. For
thoroughness, we explain how we estimated chemokine production rate
below.

There is no known literature with chemokine secretion rate
estimations.  In this paper, we present new empirical data of
\textit{in vitro} chemokine levels and extend a previously published
ODE model to estimate chemokine production rates.  We are as surprised
as Reviewer 1 that these production rates do not have a strong effect
on the agent-based modeling results (see sensitivity analyses).
Nevertheless, because our agent-based model examines chemokine
interactions, we require \textit{some} non-zero value for the production
rate.  We believe that fitting the extended ODE model to our new
empirical data is the best approach available to determine
biologically plausible values.  Our paper contains a subsection,
\textit{Models for Parameter Estimation}, that details the exact
process used to estimate the chemokine production rates (page 6).  We
fit the system of differential equations to empirical data using a
genetic algorithm to minimize the log squared error between the model
and the data (the genetic algorithm gave better fits than the more standard Levenberg-Marquardt nonlinear regression techniques).

We have now added a figure displaying the model fits to the supplement (Fig. S1, also below) and Table S3 showing the root mean squared error of the fits for completeness.

@@

  We agree that in vitro and in vivo parameters may be different, but
  this is a caveat that applies throughout biological and biomedical
  sciences.  We, along with most other researchers in the field, use
  \textit{in vitro} generated values as a starting point when
  \textit{in vivo} values are not available, and we state when we do this explicitly
  in the manuscript.  In addition, the paper contains an extensive
  sensitivity analysis to address the concern that the values used to
  parameterize the model might vary from the initial starting
  values. We do not claim that the values we estimate correspond
  exactly to  \textit{in vivo} values. Instead, we use the best fit
  estimates that are currently available as a starting point, and
  conduct the sensitivity analysis to confirm the role of individual parameters on the model.
\end{response}

@@

This question is answered directly in the paper---if diffusion of
chemokine is faster than infection spread, it does not impact our
results (Table 3, Figs. S4-S7).  In the paper, we argue that direct comparison of
viral and chemokine diffusion rates is not informative. 
Because T cells climb the chemokine gradient, a chemokine's effect is
determined by the location of its local maxima, while the viral effect
is determined by the spread and concentration of the virions
themselves.  We have edited the text to make this distinction more clear (\textit{Results: Spatial Effects}, page 10).

@@

To complement the spatial analysis, we also performed multiple
sensitivity analyses of the model's free parameters.  This included scenarios where the chemokine's diffusion rate was increased by up to two orders of magnitude (Table 2).  Each sensitivity analysis showed that the chemokine diffusion rate was an insensitive parameter in the model (Table 3 and Supplement).

@@

\begin{response}
This comment restates as fact one of the conclusions of our analysis,
so we have no disagreement with this point.
%We agree with the reviewer that the T cell secretion rate is an
%insensitive parameter in our model as determined by our multiple
%sensitivity analyses.  
However, because our agent-based model investigates the recruitment of
T cells to the focus of infection, we required a non-zero secretion
rate to introduce T cells into the model.  To select a biologically
plausible value, we fit an ODE model to empirical data in our paper
(\textit{Models for Parameter Estimation}, page 7) and similarly to
the chemokine secretion rate discussed in Point 1.  We also agree that
the recruitment of T cells to the focus of the infection is important,
a point that is discussed in the \textit{Spatial Effects} subsection (page 9-10)
and in the discussion (page 15) as follows:

@@

The reviewer asks if changing certain parameters in our model would
change the outcome for virus control. Our model does not produce the
viral control that the reviewer hypothesizes it could produce, but
we believe that reporting results that do not conform to his/her
expectations should not be confused with ignoring the request to perform
the analysis.

The term `viral control' has been used by Reviewer 1 in three different contexts.  We have explicitly addressed each point in previous responses.  In summary:

\begin{enumerate}
\item \textit{You ignored my request to investigate if virus control can be good by T cells (irrespectively of the type of virus) if chemokines diffuse faster than virus particles. Please do the simulations, and rephrase your discussion depending on the results you find.}
\begin{displayquote}
We performed these simulations numerous times as part of the
sensitivity analyses, and Response 3 of this document reviews the
different roles of viral diffusion and chemokine diffusion.  
\end{displayquote}

@@

\begin{displayquote}
In summary, our model does not attempt to include the entire immune response. It represents viral
spread using parameters that likely reflect innate control mechanisms (virus secretion, infectivity, diffusion
rate, and decay rate, see Tables 1 and 2). We appreciate the reviewer’s perspective, but our goal in this
paper is to study the effect of T cells and not to fully represent the interaction between influenza strains
and the entirety of the immune response (Introduction 3rd paragraph shown below). \\

\textit{``Specifically, we focus on the interactions between activated antigen-specific CD8 T cells, 
cytokines, and replicating influenza virus. ... Therefore, instead of developing a comprehensive
immune system model, we present a spatially explicit agent-based model (ABM) to describe T
cell interactions with chemotactic signals and a dynamically growing plaque." (page 1)} \\

We feel that careful investigation of the role of early innate control, with its many complex mechanisms, in
different influenza strains is best left to future studies.
\end{displayquote}

@@

\begin{response}
We have reviewed Oreskes et. al.  and outline how our paper is
consistent with it.  This excellent paper observes that scientists often use terms incorrectly when describing their model's veracity which leads to incorrect claims and conclusions.  Specifically, Oreskes et. al. examines four terms: `verification', `validation', `calibration', and `confirmation'.
\begin{itemize}
\item \textbf{Verification:} Oreskes et. al. states that it is not
  possible to verify a numerical model by comparing it to an
  analytical model.  We use the term `verification' once (page 15, concluding paragraph) in the context of obtaining empirical data (emphasis added).
\begin{displayquote}
``Empirical \textbf{verification} of the model's sensitive parameters (viral response to IgM, infectivity, viral decay rate, viral diffusion rate, and viral production rate) will be valuable to future studies."
\end{displayquote}
\item \textbf{Validation:} This term does not appear in our paper.
\item \textbf{Calibration:} Oreskes et. al. warn of the dangers of
  using the calibration of free parameters to emprical data as a basis
  for claiming model validity.  We use the term `calibration' only
  once (Abstract) and make no claims that calibration implies model validity (emphasis added).
\begin{displayquote}
``We \textbf{calibrate} the model using viral and chemokine secretion rates we measure \textit{in vitro} together with values taken from literature."
\end{displayquote}
\item \textbf{Confirmation:} Oreskes et. al. states that observations consistent to a theory can `confirm' a theory, but warn that confirmation cannot lead to verification.  We use the term `confirm' twice and make no claims of verification or validity (emphasis added).

@@

We agree that models can inform and guide empirical research and use our model to propose
a hypothesis that additional T cells are limited in their
effectiveness due to the spatial effects of the chemokine gradient and
T cell chemotaxis.  We cannot find a place where our paper's
conclusions are inconsistent with the contents of Oreskes et al.
\end{response}
